A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals

### 2019-08-29 paper/paper.abstract.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

Scheduling of personnel in a hospital environment is vital to improving the service provided to patients and balancing the workload assigned to clinicians. Many approaches have been tried and successfully applied to generate efficient schedules in such settings. However, due to the computational complexity of the scheduling problem in general, most approaches resort to heuristics to find a non-optimal solution in a reasonable amount of time. We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital. It mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt the formulation to a variety of clinical divisions. \\

We then conducted a case study for our approach using data from the Infectious Diseases division at St. Michael's Hospital in Toronto, Canada. We analyzed and compared the results of our approach to manually-created schedules at the hospital, and found improved adherence to departmental constraints and clinician preferences. We then used simulated data to examine the the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.

>>>>

Scheduling of personnel in a hospital environment is vital to improving the service provided to patients and balancing the workload assigned to clinicians. Many approaches have been tried and successfully applied to generate efficient schedules in such settings. However, due to the computational complexity of the scheduling problem in general, most approaches resort to heuristics to find a non-optimal solution in a reasonable amount of time. We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital. Our formulation mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt it to a variety of clinical divisions. \\

We then conducted a case study for our approach using data from the Infectious Diseases division at St. Michael's Hospital in Toronto, Canada. We analyzed and compared the results of our approach to manually-created schedules at the hospital, and found improved adherence to departmental constraints and clinician preferences. We used simulated data to examine the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.

### 2019-08-29 paper/tbl/runtime_services_clinicians.tex
Rename divisions -> services throughout
Also additional changes to results section following comments
Update runtimes in services table
Minor fixes to wording in simulations
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

\begin{table}[htbp]
	\centering
	\begin{tabular}{|c|c||c|c||c|c|}
		\hline
		                                      & \multicolumn{5}{c|}{Number of Divisions} \\ \hline
		\makecell[l]{Number of \\ Clinicians} &  1   &  2   & 2 (CBA) &  3   &  3 (CBA)  \\ \hline
		                 10                   & 0.19 & 0.95 &  0.21   & 1.90 &   0.31    \\ \hline
		                 20                   & 0.31 &  -   &  0.41   &  -   &   0.56    \\ \hline
		                 30                   & 0.57 &  -   &  0.68   &  -   &   0.90    \\ \hline
		                 50                   & 0.78 &  -   &  1.09   &  -   &   1.66    \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. `-' indicates that a solution was not found after 24 hours. `CBA' indicates that \underline{c}onsecutive \underline{b}locks were \underline{a}llowed when generating these schedules.}
	\label{tbl:runtime-divisions-comparison}%
\end{table}

>>>>

\begin{table}[htbp]
	\centering
	\begin{tabular}{|c|c||c|c||c|c|}
		\hline
		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \hline
		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \hline
		                 10                   & 0.16 &  0.74   &  0.16   & 1.40 &  0.23   \\ \hline
		                 20                   & 0.25 & 7468.86 &  0.32   &  -   &  0.43   \\ \hline
		                 30                   & 0.42 &    -    &  0.49   &  -   &  0.66   \\ \hline
		                 50                   & 0.62 &    -    &  0.82   &  -   &  1.14   \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of services and total clinicians in the division. `-' indicates that a solution was not found within 24 hours. `without NCB' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.}
	\label{tbl:runtime-services-clinicians-comparison}%
\end{table}
